---
title: "EDA nightmare analysis"
output: html_notebook
---

 

```{r, echo = FALSE}
library(tidyverse)
library(magrittr)
library(lme4)
library(lmerTest)
# READ DATA
merged <- read_csv("./data/merged_data.csv")

```

# Check normality assumption for SCR
```{r}

shapiro.test(merged$SCR)
ggplot(data = merged) +
        aes(SCR) +
        geom_histogram()

ggplot(data = merged) +
        aes(sample = SCR) +
        geom_qq()

```

The data is not normally distributed. Try sqrt transform (as SCR is always >1)

```{r}
merged <-
        merged %>%
        filter(!TestId %in% c("01", "05", "18", "31")) %>% # Removing excluded participants
        group_by(TestId) %>%
        mutate(SCR = SCR %>% sqrt %>% scale %>% as.numeric()) %>%
        ungroup()  %>%
        mutate(picture_set = if_else(session == "emot" & is.na(picture_set), "Original", picture_set))

# Check if it is better (yes)
ggplot(data = merged) +
        aes(SCR) +
        geom_histogram()

```
Much better this way.

# Build lmer with a top-down method
## 1. Disregard random structure to build a simple lm model
```{r}
nm.lm <- lm(SCR ~ group * block * category, data = merged)
summary(nm.lm)
AIC(nm.lm)

```

## 2. Use full model with full random structure (random intercept and slope)
This proves that the random structure in the data has additional value to the model.
We should select between random structures: random intercept or random intercept and slope model
The latter is the better.
```{r}
nm.lmer.is <- lmer(SCR ~ group * block * category + (block|TestId), data = merged, REML = F)
nm.lmer.i <- lmer(SCR ~ group * block * category + (1|TestId), data = merged, REML = F)

AIC(nm.lmer.is) # Best AIC, but only by a small margin
AIC(nm.lmer.i)

anova(nm.lmer.is, nm.lmer.i) # Random intercept is a better model
summary(nm.lmer.is)
```


```{r}
# 3. Removing fixed terms from the model
nm.model <- nm.lmer.is # This is the structure we go forward

# 3-way interaction
nm.3I <- update(nm.model, . ~ . - group:block:category) # Remove 3way interaction
anova(nm.model, nm.3I) # Does not add to the model

nm.model <- nm.3I
# 2-way interaction
nm.2I.1 <-  update(nm.model, . ~ . - group:block)
nm.2I.2 <-  update(nm.model, . ~ . - group:category)
nm.2I.3 <-  update(nm.model, . ~ . - block:category)

anova(nm.model, nm.2I.1)

# group:block does not improve the model, so remove
nm.model <- nm.2I.1

# Get rid of unimportant main effects
nm.M.1 <-  update(nm.model, . ~ . - group)
nm.M.2 <-  update(nm.model, . ~ . - block)
nm.M.3 <-  update(nm.model, . ~ . - category)

anova(nm.model, nm.M.3) # No main effect should be removed

```


## So the final model is:
```{r}
summary(nm.model)

merged %>% 
        filter(session == "emot") %>% 
        drop_na(block) %>% 
        group_by(block, category, group) %>% 
        summarise(Mean = mean(SCR, na.rm = T),
                  Sd = sd(SCR, na.rm = T),
                  N = n(),
                  Se = Sd/sqrt(N)) %>% 
        ggplot() +
        aes(x = block, y = Mean, ymin = Mean-Se, ymax = Mean + Se) +
        geom_errorbar(width = .2) +
        geom_col(position = "dodge") +
        facet_grid(group~category, scales = "free_x") +
        geom_smooth(method = lm, se = FALSE)

```
# Conclusion
The model shows that:  
- **Checking paradigm assumptions:** There was a habituation in responses in general (block main effect), and the habituation was larger for the Negative pictures (block x category interaction). Negative pictures were associated with larger SCR in general (category main effect).  
- **Checking main hypothesis:** The initial hypothesis of the three way introduction can not be proven, so NM participants do not show slower habilitiation than non-NM for negative stimuli. In fact the habituation of NM and NNM is the quite similar.  
- **Significant effects: ** There were differences though between NM and NNM groups. NM participants had a higher skin response to neutral stimuli in general compared to NNM (group x category). 

# Checking the recall data

```{r}
## Build lmer with a top-down method
# 1. Disregard random structure 
morning.lm <- lm(SCR ~ group * category * session + picture_set*group*category, data = merged)
morning.i <- lmer(SCR ~ group * category * session + picture_set*group*category + (1|TestId), data = merged)
morning.is <- lmer(SCR ~ group * category * session + picture_set*group*category + (session|TestId), data = merged) # We don't have 4-way interaction in the model

# The best model is the random intercept, random slope model
AIC(morning.lm)
anova(morning.i, morning.is)
morning.model <- morning.is
summary(morning.model)
```

```{r}
## Removing interacions
# Checking 3-way interactions
morning.3I.1 <- update(morning.model, . ~ . - group:session:category) # Drop
morning.3I.2 <- update(morning.model, . ~ . - group:picture_set:category) # Drop

anova(morning.model, morning.3I.1, morning.3I.2)
morning.model <- update(morning.model, . ~ . -group:session:category -group:picture_set:category ) 
```

```{r}
# Checking 2-way interactions
morning.2I.1 <- update(morning.model, . ~ . - group:session) # Drop
morning.2I.2 <- update(morning.model, . ~ . - picture_set:category) # Drop
morning.2I.3 <- update(morning.model, . ~ . - group:category) # Do not drop
morning.2I.4 <- update(morning.model, . ~ . - group:picture_set) # Drop
morning.2I.5 <- update(morning.model, . ~ . - session:category) # Drop

morning.model <- update(morning.model, . ~ . - group:session - group:picture_set - picture_set:category - session:category) 
morning.model %>% summary
```

```{r}
# Checking main effects
morning.M.1 <- update(morning.model, . ~ . - group) 
morning.M.2 <- update(morning.model, . ~ . - session) # Dop
morning.M.3 <- update(morning.model, . ~ . - picture_set) # Drop
morning.M.4 <- update(morning.model, . ~ . - category) 

anova(morning.model, morning.M.1, morning.M.2, morning.M.3, morning.M.4)
morning.model <- update(morning.model, .~. - session -picture_set)
```

```{r}
# Final model
morning.model %>% summary

```

```{r}
# Plot the aggregated SCR response to the stimuli for each day, 
merged %>% 
        drop_na(picture_set, category) %>% 
        mutate(picture_set = picture_set %>% factor(levels = c("Original", "New")),
               ) %>% 
        group_by(picture_set, session, category, group) %>% 
        summarise(Mean = mean(SCR, na.rm = T),
                  Sd = sd(SCR, na.rm = T),
                  N = n(),
                  Se = Sd/sqrt(N)) %>% 
        ggplot() +
        aes(x = session, y = Mean, ymin = Mean-Se, ymax = Mean + Se, fill = picture_set, group = picture_set) +
        geom_errorbar(position = position_dodge()) +
        geom_col(position = position_dodge()) +
        facet_grid(group ~ category)
```
```{r}
merged %>% 
        drop_na(picture_set, category) %>% 
        mutate(picture_set = picture_set %>% factor(levels = c("Original", "New")),
        ) %>% 
        group_by(picture_set, session, category, group) %>% 
        summarise(Mean = mean(SCR, na.rm = T),
                  Sd = sd(SCR, na.rm = T),
                  N = n(),
                  Se = Sd/sqrt(N)) %>% 
        ggplot() +
        aes(x = group, y = Mean, ymin = Mean-Se, ymax = Mean + Se, fill = picture_set, group = picture_set) +
        geom_errorbar(position = position_dodge()) +
        geom_col(position = position_dodge()) +
        facet_grid(session ~ category)
```
# Conclusion
The model yields that:  

- Negative pictures elicit larger SCR response than neutral (category main effect).
- The nightmare group showed a smaller SCR reaction to negative pictures, but they also gave a larger SCR response to neutral pictures (group x category interaction)